A SIMPLE PRIMAL-DUAL ALGORITHM FOR COMPUTING APPROXIMATE NASH-EQUAILIBRIA IN SEQUENTIAL GAMES WITH INCOMPLETE INFORMATION

### 2015-10-02 paper.tex
fixups

\subsection{Quick sketch of our contribution}
%% One cannot directly attack the LCPs
%% \eqref{eq:primal_pb} and
%% \eqref{eq:dual_pb} via a traditional primal-dual algorithm
%% (for example \cite{chambolle2010,chambolle2014ergodic}) because
%% computing the orthogonal projections $\proj_{Q_k}$ is very difficult
%% In fact, such sub-problems would have to be solved
%% iteratively\footnote{An ``exception to the rule'' is the case where
%%   the $Q_k's$ are simplexes, so that the projections can be computed
%%   exactly using \cite{duchi2008efficient}.}.

Developing on an alternative notion of approximate equilibria (see
Definition \ref{thm:cool_notion})
homologous to that presented in Definition \ref{thm:approx_nash}, we
device a simple numerically stable primal-dual algorithm that
(Algorithm \ref{Tab:algo}) for computing approximate Nash-equilibria
in sequential two-person zero-sum games with incomplete information and
perfect recall. On, each iteration, the only operations performed by
our algorithm are of the form $A^Tx$, $Ay$, $E_1^Tp$, $E_2^Tq$, and
$(x)_+$. We also prove (Theorem \ref{thm:pd}) that --in an ergodic /
Ces\`ario sense-- the number of iterations required by the algorithm
to produce an approximation equilibrium to a precision $\epsilon$ is
$\mathcal{O}(1/\epsilon)$, with explicit values for the constants
involved this worst-case cost.
%% The source of approximation in our scheme
%% is that we explicitly control how close we are to the feasible set
%% $Q_1 \times Q_2$.
% These contributions will be elaborated in section \ref{sec:gsp}.

\section{Relation to prior work}
\label{sec:related_work}
We now present a selection of algorithms that is representative of the
efforts that have been made in the literature to compute Nash
$\epsilon$-equilibria for two-person zero-sum games with incomplete
information like Texas Hold'em Poker, etc.
First and foremost, let us note that for the class of games considered
here (sequential games with incomplete information), the LCPs
\eqref{eq:primal_pb} and \eqref{eq:dual_pb} are exceedingly larger
than what state-of-the-art LCP and interior-point solvers can
handle (see for example \cite{hoda2010smoothing,gilpinfirst}).

\subsection{Regret minimization}
CFR (CounterFactual Regret minimization) \cite{zinkevich2008regret},
Monte Carlo CFR \cite{lanctot2009monte}, and CFR+
\cite{Bowling09012015}, by their large popularity, have become the
definitive state-of-the-art, and are particularly useful in
many-player games, since convex-analytical methods cannot help much in
such games. Also, they can be shown to converge to a Nash-equilibrium
provided each player uses a CFR scheme to play the game
\cite{zinkevich2008regret}, but generally enjoy a much weaker
convergence theory. For example, \cite{lanctot2009monte} showed that
such scheme have a running time of $\mathcal{O}(1/\epsilon^2)$ to
produce a Nash $\epsilon$-equilibrium.

@@

\section{Our contributions}
\label{sec:gsp}
\subsection{Preliminaries}
In the next theorem, we show that the LCPs \eqref{eq:primal_pb}
and \eqref{eq:dual_pb} can be
conveniently written as a \textit{Generalized Saddle-point Problem
  (GSP)} in the sense of \cite{he2013accelerating}. The crux of idea
is to remove the linear constraints in the definitions of the strategy
polytopes $Q_k$, by augmenting the payoff matrix to yield an equivalent
saddle-point problem.
\begin{theorem}
Define two proper closed convex functions
  \begin{eqnarray}
    \left.
    \begin{aligned}
      g_1: \mathbb{R}^{n_2} &\times \mathbb{R}^{l_1} \rightarrow
      (-\infty, +\infty], \hspace{1em} g_1(y, p) :=
        i_{y \ge 0} + \langle e_1,p\rangle\\
        g_2: \mathbb{R}^{n_1} &\times \mathbb{R}^{l_2} \rightarrow
        (-\infty, +\infty],\hspace{1em} g_2(x, q) :=
          i_{x \ge 0} + \langle e_2, q\rangle
    \end{aligned}
    \right\}
    \label{eq:things}
  \end{eqnarray}
Also define two bilinear forms $\Psi_1$, $\Psi_2: \mathbb{R}^{n_2}
\times \mathbb{R}^{l_1} \times \mathbb{R}^{n_1} \times
\mathbb{R}^{l_2} \rightarrow \mathbb{R}$ with $\Psi_2 = -\Psi_1$ by letting
\begin{equation}
  \begin{split}  
    K :=
    \left[
      \begin{array}{cc}
        A & -E_1^T \\
        E_2 & 0
      \end{array}
      \right],\hspace{.5em}
    \Psi_1(y, p, x, q)
    := \left\langle \begin{bmatrix}x\\q\end{bmatrix},
      K\begin{bmatrix}y\\p\end{bmatrix}\right\rangle%% \\
      %% &= \langle x,
      %% Ay\rangle -\langle x, E_1^Tp\rangle + \langle q, E_2y\rangle,
\end{split}
\end{equation}
and define the functions $\hat{\Psi}_1$, $\hat{\Psi}_2:
\mathbb{R}^{n_2} \times \mathbb{R}^{l_1} \times \mathbb{R}^{n_1}
\times \mathbb{R}^{l_2} \rightarrow (-\infty, +\infty]$ by
\begin{eqnarray}
  \begin{aligned}
    \hat{\Psi}_1(y, p, x, q) :=\begin{cases}
    \Psi_1(y, p, x, q)+ g_1(y, p), &\mbox{ if }y \ge 0,\\
    +\infty, &\mbox{ otherwise}\end{cases}\\
    %% &\hat{\Psi}_2(y, p, x, q) &:= \Psi_2(y, p, x, q)+ g_2(x, q) \text{ if
    %% }x \ge 0, \hat{\Psi}_2(y,p, x, q) := \infty \text{ otherwise}.
    \hat{\Psi}_2(y, p, x, q) :=\begin{cases}
    \Psi_2(y, p, x, q)+ g_2(x, q), &\mbox{ if }x \ge 0,\\
    +\infty, &\mbox{ otherwise}\end{cases}
  \end{aligned}
\end{eqnarray}
Finally, define the sets $S_1 := \mathbb{R}^{n_2}_+ \times
\mathbb{R}^{l_1}$ and $S_2 := \mathbb{R}^{n_1}_+ \times
\mathbb{R}^{l_2}$, and consider the GSP($\Psi_1$, $\Psi_2$, $g_1$,
$g_2$): Find a quadruplet $(y^*,p^*, x^*, q^*) \in S_1 \times S_2$,
\begin{eqnarray}
  \begin{split}
    &\hat{\Psi}_1(y^*, p^*, x^*, q^*) \le \hat{\Psi}_1(y, p, x^*,
    q^*),\hspace{1em}and\\
    &\hat{\Psi}_2(y^*, p^*, x^*, q^*)
    \le \hat{\Psi}_2(y^*, p^*, x, q),\hspace{.21em} \forall (y,p, x, q)
    \in S_1 \times S_2.
  \label{eq:unconstrained_pb}
\end{split}
\end{eqnarray}
\label{thm:pd}
Then GSP($\Psi_1$,
  $\Psi_2$, $g_1$, $g_2$) is equivalent to the LCPs
  \eqref{eq:primal_pb} and \eqref{eq:dual_pb}, i.e
a quadruplet $(y^*,p^*, x^*, q^*) \in \mathbb{R}^{n_2}
  \times \mathbb{R}^{l_1} \times \mathbb{R}^{n_1} \times
  \mathbb{R}^{l_2}$ solves the LCPs
  \eqref{eq:primal_pb} and \eqref{eq:dual_pb} iff it solves
  GSP($\Psi_1$, $\Psi_2$, $g_1$, $g_2$). 
  \label{thm:pd}
\end{theorem}

@@

\paragraph*{Computing $\|K\|$.}
Also the 2-norm $\|K\|$ of the linear operator $K$, can be efficiently
computed using the power iteration. Results are presented and
commented in Figure \ref{Tab:dgap_curve}.

\section{Numerical experiments results}
\label{sec:results}
We now present some proof-of-concept for the algorithm proposed.
\subsection{Basic test-bed: Matrix games on simplexes}
As in \cite{nesterov2005a,chambolle2014ergodic}, we generate a $1000
\times 1000$ random matrix whose entries are uniformly identically
distributed in the closed interval $[-1, 1]$. %%  We compare our proposed
%% Algorithm \ref{Tab:algo} with Nesterov's \cite{nesterov2005a} and
%% Gilpin's et al. \cite{gilpinfirst}.
The results of the experiments are shown in Figure
\ref{Tab:dgap_curve}\textit{(a)}.

>>>>

\subsection{Quick sketch of our contribution}
%% One cannot directly attack the LCPs
%% \eqref{eq:primal_pb} and
%% \eqref{eq:dual_pb} via a traditional primal-dual algorithm
%% (for example \cite{chambolle2010,chambolle2014ergodic}) because
%% computing the orthogonal projections $\proj_{Q_k}$ is very difficult
%% In fact, such sub-problems would have to be solved
%% iteratively\footnote{An ``exception to the rule'' is the case where
%%   the $Q_k's$ are simplexes, so that the projections can be computed
%%   exactly using \cite{duchi2008efficient}.}.
We now give a brief overview of our contributions, which will be made
more elaborate in section \ref{sec:gsp}.
Developing on an alternative notion of approximate equilibria (see
Definition \ref{thm:cool_notion})
homologous to that presented in Definition \ref{thm:approx_nash}, we
device a simple numerically stable primal-dual algorithm that
(Algorithm \ref{Tab:algo}) for computing approximate Nash-equilibria
in sequential two-person zero-sum games with incomplete information and
perfect recall. On, each iteration, the only operations performed by
our algorithm are of the form $A^Tx$, $Ay$, $E_1^Tp$, $E_2^Tq$, and
$(x)_+$. We also prove (Theorem \ref{thm:pd}) that --in an ergodic /
Ces\`ario sense-- the number of iterations required by the algorithm
to produce an approximation equilibrium to a precision $\epsilon$ is
$\mathcal{O}(1/\epsilon)$, with explicit values for the constants
involved.
%% The source of approximation in our scheme
%% is that we explicitly control how close we are to the feasible set
%% $Q_1 \times Q_2$.
% These contributions will be elaborated in section \ref{sec:gsp}.

\section{Relation to prior work}
\label{sec:related_work}
Here, we present a selection of algorithms that is representative of the
efforts that have been made in the literature to compute Nash
$\epsilon$-equilibria for two-person zero-sum games with incomplete
information like Texas Hold'em Poker, etc.
First and foremost, let us note that for the class of games considered
here (sequential games with incomplete information), the LCPs
\eqref{eq:primal_pb} and \eqref{eq:dual_pb} are exceedingly larger
than what state-of-the-art LCP and interior-point solvers can
handle (see for example \cite{hoda2010smoothing,gilpinfirst}).

\subsection{Regret minimization}
CFR (CounterFactual Regret minimization) \cite{zinkevich2008regret},
Monte Carlo CFR \cite{lanctot2009monte}, and CFR+
\cite{Bowling09012015}, by their large popularity, have become the
definitive state-of-the-art, and are particularly useful in
many-player games, since convex-analytical methods cannot help much in
such games. Also, they can be shown to converge to a Nash-equilibrium
provided each player uses a CFR scheme to play the game
\cite{zinkevich2008regret}, but have a much weaker
convergence theory. For example, \cite{lanctot2009monte} showed that
such scheme have a running time of $\mathcal{O}(1/\epsilon^2)$ to
produce a Nash $\epsilon$-equilibrium.

@@

\section{Our contributions}
\label{sec:gsp}
\subsection{Preliminaries}
In the next theorem, we show that the LCPs \eqref{eq:primal_pb}
and \eqref{eq:dual_pb} can be
conveniently written as a \textit{Generalized Saddle-point Problem
  (GSP)} in the sense of \cite{he2013accelerating}. The crux of idea
is to remove the linear constraints in the definitions of the strategy
polytopes $Q_k$, by augmenting the payoff matrix to yield an equivalent
saddle-point problem.
\begin{theorem}
Define two proper closed convex functions
  \begin{eqnarray}
    \left.
    \begin{aligned}
      g_1: \mathbb{R}^{n_2} &\times \mathbb{R}^{l_1} \rightarrow
      (-\infty, +\infty], \hspace{1em} g_1(y, p) :=
        i_{y \ge 0} + \langle e_1,p\rangle\\
        g_2: \mathbb{R}^{n_1} &\times \mathbb{R}^{l_2} \rightarrow
        (-\infty, +\infty],\hspace{1em} g_2(x, q) :=
          i_{x \ge 0} + \langle e_2, q\rangle
    \end{aligned}
    \right\}
    \label{eq:things}
  \end{eqnarray}
Also define two bilinear forms $\Psi_1$, $\Psi_2: \mathbb{R}^{n_2}
\times \mathbb{R}^{l_1} \times \mathbb{R}^{n_1} \times
\mathbb{R}^{l_2} \rightarrow \mathbb{R}$ with $\Psi_2 = -\Psi_1$ by letting
\begin{equation}
  \begin{split}  
    K :=
    \left[
      \begin{array}{cc}
        A & -E_1^T \\
        E_2 & 0
      \end{array}
      \right],\hspace{.5em}
    \Psi_1(y, p, x, q)
    := \left\langle \begin{bmatrix}x\\q\end{bmatrix},
      K\begin{bmatrix}y\\p\end{bmatrix}\right\rangle%% \\
      %% &= \langle x,
      %% Ay\rangle -\langle x, E_1^Tp\rangle + \langle q, E_2y\rangle,
\end{split}
\end{equation}
and define the functions $\hat{\Psi}_1$, $\hat{\Psi}_2:
\mathbb{R}^{n_2} \times \mathbb{R}^{l_1} \times \mathbb{R}^{n_1}
\times \mathbb{R}^{l_2} \rightarrow (-\infty, +\infty]$ by
\begin{eqnarray}
  \begin{aligned}
    \hat{\Psi}_1(y, p, x, q) :=\begin{cases}
    \Psi_1(y, p, x, q)+ g_1(y, p), &\mbox{ if }y \ge 0,\\
    +\infty, &\mbox{ otherwise}\end{cases}\\
    %% &\hat{\Psi}_2(y, p, x, q) &:= \Psi_2(y, p, x, q)+ g_2(x, q) \text{ if
    %% }x \ge 0, \hat{\Psi}_2(y,p, x, q) := \infty \text{ otherwise}.
    \hat{\Psi}_2(y, p, x, q) :=\begin{cases}
    \Psi_2(y, p, x, q)+ g_2(x, q), &\mbox{ if }x \ge 0,\\
    +\infty, &\mbox{ otherwise}\end{cases}
  \end{aligned}
\end{eqnarray}
Finally, define the sets $S_1 := \mathbb{R}^{n_2}_+ \times
\mathbb{R}^{l_1}$ and $S_2 := \mathbb{R}^{n_1}_+ \times
\mathbb{R}^{l_2}$, and consider the GSP($\Psi_1$, $\Psi_2$, $g_1$,
$g_2$): Find a quadruplet $(y^*,p^*, x^*, q^*) \in S_1 \times S_2$
such that
\begin{eqnarray}
  \begin{split}
    &\hat{\Psi}_1(y^*, p^*, x^*, q^*) \le \hat{\Psi}_1(y, p, x^*,
    q^*),\hspace{1em}and\\
    &\hat{\Psi}_2(y^*, p^*, x^*, q^*)
    \le \hat{\Psi}_2(y^*, p^*, x, q),\hspace{.21em} \forall (y,p, x, q)
    \in S_1 \times S_2.
  \label{eq:unconstrained_pb}
\end{split}
\end{eqnarray}
\label{thm:pd}
Then GSP($\Psi_1$,
  $\Psi_2$, $g_1$, $g_2$) is equivalent to the LCPs
  \eqref{eq:primal_pb} and \eqref{eq:dual_pb}, i.e
a quadruplet $(y^*,p^*, x^*, q^*) \in \mathbb{R}^{n_2}
  \times \mathbb{R}^{l_1} \times \mathbb{R}^{n_1} \times
  \mathbb{R}^{l_2}$ solves the LCPs
  \eqref{eq:primal_pb} and \eqref{eq:dual_pb} iff it solves
  GSP($\Psi_1$, $\Psi_2$, $g_1$, $g_2$). 
  \label{thm:pd}
\end{theorem}

@@

\paragraph*{Computing $\|K\|$.}
Also the 2-norm $\|K\|$ of the linear operator $K$, can be efficiently
computed using the power iteration.

\section{Numerical experiments results}
\label{sec:results}
We now present some proof-of-concept for the algorithm proposed.
Results are presented and commented in Figure \ref{Tab:dgap_curve}.
\subsection{Basic test-bed: Matrix games on simplexes}
As in \cite{nesterov2005a,chambolle2014ergodic}, we generate a $1000
\times 1000$ random matrix whose entries are uniformly identically
distributed in the closed interval $[-1, 1]$. %%  We compare our proposed
%% Algorithm \ref{Tab:algo} with Nesterov's \cite{nesterov2005a} and
%% Gilpin's et al. \cite{gilpinfirst}.
The results of the experiments are shown in Figure
\ref{Tab:dgap_curve}\textit{(a)}.
